---
title: "fxTWAPLS2 climate reconstruction"
output: html_document
---


```{r}
#You can install the released version of fxTWAPLS from CRAN with:
install.packages("fxTWAPLS")
```


#Input data
1. Modern pollen for trainning the model
2. Fossil pollen
```{r}
modern_pollen <- read.csv("Input/train_set_amg.csv")
fossil_pollen <- read.csv("Input/pollen_proportions.csv")
```


#Obtain modern pollen taxa
```{r}
#the start and end taxa are Abies and Zygophyllaceae in the modern pollen dataset
taxaColMin <- which(colnames(modern_pollen) == "Abies")
taxaColMax <- which(colnames(modern_pollen) == "Zygophyllaceae")
taxa <- modern_pollen[, taxaColMin:taxaColMax]
```


# Use fxTWAPLSv2 to train
```{r}
# MTCO
fit_tf_Tmin <- fxTWAPLS::TWAPLS.w2(taxa, modern_pollen$Tmin, nPLS = 5, usefx = TRUE, fx_method="pspline",bin=0.02)

# GDD0
fit_tf_gdd <- fxTWAPLS::TWAPLS.w2(taxa, modern_pollen$gdd, nPLS = 5, usefx = TRUE, fx_method="pspline",bin=20)

# alpha
fit_tf_alpha <- fxTWAPLS::TWAPLS.w2(taxa, modern_pollen$alpha, nPLS = 5, usefx = TRUE, fx_method="pspline",bin=0.002)

#MTWA
fit_tf_Tmax <- fxTWAPLS::TWAPLS.w2(taxa, modern_pollen$Tmax, nPLS = 5, usefx = TRUE, fx_method="pspline",bin=0.02)
```


# Cross validation (pseudo removed leave out cross validation)
# Get the distance between each point
```{r}
# Get the location information of each sample
point <- modern_pollen[, c("Long", "Lat")]
# Get the distance between each point
dist <- fxTWAPLS::get_distance(point, cpus = 8)
# Set CPUS to run in parallel
CPUS<-4

# Get the pseudo sites (which are both geographically close and climatically
# close to the test site) which should be removed in cross validation
pseudo_Tmin <- fxTWAPLS::get_pseudo(dist, modern_pollen$Tmin, cpus = CPUS)
pseudo_gdd <- fxTWAPLS::get_pseudo(dist, modern_pollen$gdd, cpus = CPUS)
pseudo_alpha <- fxTWAPLS::get_pseudo(dist, modern_pollen$alpha, cpus = CPUS)
pseudo_Tmax <- fxTWAPLS::get_pseudo(dist, modern_pollen$Tmax, cpus = CPUS)

rlist::list.save(pseudo_Tmin, 'pseudo_Tmin.rdata')
rlist::list.save(pseudo_gdd, 'pseudo_gdd.rdata')
rlist::list.save(pseudo_alpha, 'pseudo_alpha.rdata')
rlist::list.save(pseudo_Tmax, 'pseudo_Tmax.rdata')


pseudo_Tmin <- rlist::list.load('pseudo_Tmin.rdata')
pseudo_gdd <- rlist::list.load('pseudo_gdd.rdata')
pseudo_alpha <- rlist::list.load('pseudo_alpha.rdata')
pseudo_Tmax <- rlist::list.load('pseudo_Tmax.rdata')


# Leave-out cross validation
`%>%` <- magrittr::`%>%`
if(!require(foreach)){install.packages("foreach");library(foreach)}
cv_tf2_Tmin_pspline <- fxTWAPLS::cv.pr.w(taxa,
                                         modern_pollen$Tmin,
                                         nPLS = 5,
                                         fxTWAPLS::TWAPLS.w2,
                                         fxTWAPLS::TWAPLS.predict.w,
                                         pseudo_Tmin,
                                         usefx = TRUE,
                                         fx_method = "pspline",
                                         bin = 0.02,
                                         cpus = 8,
                                         test_mode = F)  %>% fxTWAPLS::pb()  
write.csv(cv_tf2_Tmin_pspline, "Output_train_amg/cv_tf2_Tmin_pspline.csv")

cv_tf2_gdd_pspline <- fxTWAPLS::cv.pr.w(taxa,
                                        modern_pollen$gdd,
                                        nPLS = 5,
                                        fxTWAPLS::TWAPLS.w2,
                                        fxTWAPLS::TWAPLS.predict.w,
                                        pseudo_gdd,
                                        usefx = TRUE,
                                        fx_method = "pspline",
                                        bin = 20,
                                        cpus = 8,
                                        test_mode = F)   %>% fxTWAPLS::pb()  
write.csv(cv_tf2_gdd_pspline, "Output_train_amg/cv_tf2_gdd_pspline.csv")

cv_tf2_alpha_pspline <- fxTWAPLS::cv.pr.w(taxa,
                                          modern_pollen$alpha,
                                          nPLS = 5,
                                          fxTWAPLS::TWAPLS.w2,
                                          fxTWAPLS::TWAPLS.predict.w,
                                          pseudo_alpha,
                                          usefx = TRUE,
                                          fx_method = "pspline",
                                          bin = 0.002,
                                          cpus = 8,
                                          test_mode = F)   %>% fxTWAPLS::pb()  
write.csv(cv_tf2_alpha_pspline, "Output_train_amg/cv_tf2_alpha_pspline.csv")

cv_tf2_Tmax_pspline <- fxTWAPLS::cv.pr.w(taxa,
                                         modern_pollen$Tmax,
                                         nPLS = 5,
                                         fxTWAPLS::TWAPLS.w2,
                                         fxTWAPLS::TWAPLS.predict.w,
                                         pseudo_Tmax,
                                         usefx = TRUE,
                                         fx_method = "pspline",
                                         bin = 0.02,
                                         cpus = 8,
                                         test_mode = F)   %>% fxTWAPLS::pb()  
write.csv(cv_tf2_Tmax_pspline, "Output_train_amg/cv_tf2_Tmax_pspline.csv")

# Random t test to the cross validation result
rand_tf2_Tmin_pspline <- fxTWAPLS::rand.t.test.w(cv_tf2_Tmin_pspline, n.perm = 999) 
rand_tf2_alpha_pspline <- fxTWAPLS::rand.t.test.w(cv_tf2_alpha_pspline, n.perm = 999) 
rand_tf2_Tmax_pspline <- fxTWAPLS::rand.t.test.w(cv_tf2_Tmax_pspline, n.perm = 999) 
rand_tf2_gdd_pspline <- fxTWAPLS::rand.t.test.w(cv_tf2_gdd_pspline, n.perm = 999) 
```




#Output table from cross validation results.
In EMBSeCBIO case the last last significant number of components (the number of components to be usedis 4 for MTCO, 2 for GDD0, 3 for alpha, and 4 for MTWA.
```{r}
rand_tf2_pspline<-rbind.data.frame(rand_tf2_Tmin_pspline,rand_tf2_Tmax_pspline,rand_tf2_alpha_pspline,rand_tf2_gdd_pspline)# |> mutate(n_comp=rep(seq(1,5,1),times=4))

write.csv(rand_tf2_pspline, "Output_train_amg/rand_tf2_pspline.csv")
```



# Reconstruction
```{r}
#Extract fossil taxa, the start and end taxa are Abies and Zygophyllaceae in my fossil dataset
taxaColMin <- which(colnames(fossil_pollen) == "Abies")
taxaColMax <- which(colnames(fossil_pollen) == "Zygophyllaceae")
core0<-fossil_pollen[,taxaColMin:taxaColMax]
core0[is.na(core0)]<-0;str(core0)
core<-core0

#get common taxa
colnames(core)[!(colnames(core)%in% colnames(taxa))] #taxa that are in fossil data but not exist in modern data
colnames(taxa)[!(colnames(taxa)%in% colnames(core))] #taxa that are in modern data but not exist in fossil data

colnames(core)[(colnames(core)%in% colnames(taxa))]
```



```{r}
#delete taxa that are in fossil data but not exist in modern data
taxa_to_delete<-colnames(core)[!(colnames(core)%in% colnames(taxa))]
core[,taxa_to_delete]<-NULL
#add taxa that are in modern data but not exist in fossil data and attribute them to 0
taxa_to_add<-colnames(taxa)[!(colnames(taxa)%in% colnames(core))]
core[,taxa_to_add]<-0
core<-core[,order(colnames(core))]

#recalculate the relative abundance
summary(rowSums(core))
core<-core/rowSums(core)
summary(rowSums(core))
str(core)
```



#Reconstruct the past climates
```{r}
#MTCO
fossil_tf_Tmin<-fxTWAPLS::TWAPLS.predict.w(fit_tf_Tmin,core)
#GDD0
fossil_tf_gdd<-fxTWAPLS::TWAPLS.predict.w(fit_tf_gdd,core)
#alpha
fossil_tf_alpha<-fxTWAPLS::TWAPLS.predict.w(fit_tf_alpha,core)
#MTWA
fossil_tf_Tmax<-fxTWAPLS::TWAPLS.predict.w(fit_tf_Tmax,core)

#Use the last significant number of components
core_sig<-cbind.data.frame(fossil_pollen[,c("entity_name","latitude","longitude","elevation","age")],
                           fossil_tf_Tmin[["fit"]][,4],fossil_tf_gdd[["fit"]][,2],fossil_tf_alpha[["fit"]][,2], fossil_tf_Tmax[["fit"]][,2])

colnames(core_sig)<-c("site","lat","lon","elv","age","Tmin_pred","gdd_pred","alpha_pred","Tmax_pred")

write.csv(core_sig,"Output_train_amg/Train_amg_fossil_reconstruction_without_sse.csv")


# Get the sample specific errors

`%>%` <- magrittr::`%>%`
#MTCO
sse_tf_Tmin<-fxTWAPLS::sse.sample(modern_taxa=taxa,modern_climate=modern_pollen$Tmin,fossil_taxa=core,trainfun=fxTWAPLS::TWAPLS.w2,predictfun=fxTWAPLS::TWAPLS.predict.w,
                                  nboot=1000,nPLS=5,nsig=4,usefx=TRUE,fx_method = "pspline",bin=0.02)%>%fxTWAPLS::pb()
#GDD0
sse_tf_gdd<-fxTWAPLS::sse.sample(modern_taxa=taxa,modern_climate=modern_pollen$gdd,fossil_taxa=core,trainfun=fxTWAPLS::TWAPLS.w2,predictfun=fxTWAPLS::TWAPLS.predict.w,
                                 nboot=1000,nPLS=5,nsig=2,usefx=TRUE,fx_method = "pspline",bin=20)%>%fxTWAPLS::pb()
#alpha
sse_tf_alpha<-fxTWAPLS::sse.sample(modern_taxa=taxa,modern_climate=modern_pollen$alpha,fossil_taxa=core,trainfun=fxTWAPLS::TWAPLS.w2,predictfun=fxTWAPLS::TWAPLS.predict.w,
                                   nboot=1000,nPLS=5,nsig=2,usefx=TRUE,fx_method = "pspline",bin=0.002)%>%fxTWAPLS::pb()
#MTWA
sse_tf_Tmax<-fxTWAPLS::sse.sample(modern_taxa=taxa,modern_climate=modern_pollen$Tmax,fossil_taxa=core,trainfun=fxTWAPLS::TWAPLS.w2,predictfun=fxTWAPLS::TWAPLS.predict.w,
                                  nboot=1000,nPLS=5,nsig=2,usefx=TRUE,fx_method = "pspline",bin=0.02)%>%fxTWAPLS::pb()



# Final output
sse_core_sig<-cbind.data.frame(sse_tf_Tmin,sse_tf_gdd,sse_tf_alpha,sse_tf_Tmax)
colnames(sse_core_sig)<-c("Tmin_sse","gdd_sse","alpha_see","Tmax_see")

reconstruction_output <- dplyr::bind_cols(core_sig,sse_core_sig)

write.csv(reconstruction_output,"Output_train_amg/Train_amg_reconstruction_output.csv")
```

